---
title: "Download Past SOA Exams"
author: "Shanghao Zhong"
date: "10/20/2021"
output: html_document
---

# Load packages
Install required packages if they are not installed.
```{r}
require.packages <- c('rvest', 'tidyverse')
install.packages(require.packages[!is.element(require.packages, installed.packages())])
rm(require.packages)
```

Load packages.
```{r}
library(rvest)
library(tidyverse)
```

# Web Scraping
```{r}
url <- 'https://www.soa.org/education/exam-req/syllabus-study-materials/edu-multiple-choice-exam'
doc <- read_html(url)
```

## Get all the exam season
```{r}
seasons <- doc %>% html_nodes(xpath = "////a[starts-with(@href, '/globalassets/assets/files/edu/')]/ancestor::tbody")
seasons.names <- seasons %>% map(~{.x %>% html_nodes(xpath = 'tr[1]/td[1]') %>% html_text() %>% str_extract('\\w+\\s\\d{4}')})
```

Get all the exams in a season.
```{r}
ifnull <- function(string, string2) {
  # if `string` is null, return `string2`, else return `string` instead
  if(identical(string, character(0)) || is.na(string)) {
    return(string2)
  } else {
    return(string)
  }
}

getExamName <- function(season.node) {
  # get the exam name from one of the node in `seasons`. 
  
  # since SOA's website has inconsistent structure, it will try a couple things to get the exam name.
  td.nodes <- season.node %>% html_nodes(xpath = "descendant-or-self::a/ancestor::td")
  exams <- td.nodes %>% map(~{
      .x %>% html_nodes('strong') %>% html_text() %>% ifnull(
        .x %>% html_nodes(xpath = 'ul/li[1]/a') %>% html_attr('title') %>% ifnull(
          .x %>% html_nodes(xpath = 'ul/li[1]/a') %>% html_text()
        )
      )
    }) %>% unlist()
  
  # get exam versions, like exam dates.
  versions <- season.node %>% 
    html_nodes(xpath = "descendant-or-self::a/ancestor::tbody/tr[2]/th") %>% 
    html_text() %>% str_extract('\\w* \\d*, \\d{4}') %>% 
    parse_date("%B %d, %Y") %>% format("%m%d")
  
  if(length(versions) == 0) {
    versions <- c(NA)
  }
  
  return(data.frame(exam = exams, version = versions))
}

seasons.exams <- seasons %>% map(~getExamName(.x)) 
```

Create a list of exams.
```{r}
exam.list <- map2_dfr(seasons.names, seasons.exams, ~{data.frame(season = .x, .y)})
```

Do some clean up.
```{r}
exam.data <- exam.list %>%
  mutate(year = str_extract(season, "\\d{4}") %>% parse_integer(),
         season = str_extract(season, "\\w+"),
         version = coalesce(version, str_extract(exam, "(Morning)|(Afternoon)")),
         exam = exam %>% str_replace('.*ERM.*', 'Exam ERM') %>% str_extract("Exam \\w+(( |–)\\w+)*")) %>%
  relocate(year, .before = season)
```

## Get all the file links
Get all the nodes with file links.
```{r}
hrefs <- doc %>% html_nodes(xpath = "////a[starts-with(@href, '/globalassets/assets/files/edu/')]/ancestor::td")
```

Get all the files and links.
```{r}
links <- hrefs %>% map(~{
  data.frame(file=.x %>% html_nodes(xpath = 'descendant-or-self::a') %>% html_text(),
             link=.x %>% html_nodes(xpath = 'descendant-or-self::a') %>% html_attr('href'))
})
```

Create a data frame of all files.
```{r}
files.data <- map2_dfr(
  split(exam.data, seq(nrow(exam.data))), 
  links, 
  ~{data.frame(.x %>% `rownames<-`(NULL), .y)})
```

Remove intermediate variables.
```{r}
rm(list = setdiff(ls(), 'files.data'))
```

# Download Files
Create a function that download all files in the `df`.
```{r}
download.files <- function(df) {
  for(i in seq(nrow(df))) {
    link <- str_glue('https://www.soa.org', df[i, 'link'])
    folder <- str_glue(df[i, 'exam'] %>% str_replace(' ', '-'), df[i, 'year'], df[i, 'season'], .sep = "_")
    if(!is.na(df[i, 'version'])) {
      folder <- str_glue(folder, '_', df[i, 'version'])
    }
    file.name <- link %>% str_split('/') %>% unlist() %>% last()
    path <- file.path(folder, file.name)
    
    # create folder
    dir.create(folder, showWarnings = FALSE)
    
    # download file
    download.file(link, path)
  }
}
```

## What you can download
Here are all the available exams.
```{r}
files.data %>% select(exam) %>% unique() %>% arrange(exam)
```

You can also see all the available exam sittings. (It will open a new tab in RStudio.)
```{r}
files.data %>% select(year, season, exam, version) %>% unique() %>% View()
```

## Download the files
Here, it downloads all the files for Exam PA. You can also change the code to download the things you want. (It will open a new tab in RStudio.)
```{r}
download.files.data <- files.data %>% filter(exam == 'Exam PA')
View(download.files.data)
```

Use `download.files` to get all the exam files.
Files will be downloaded to the same location as this R project. 
There will be one folder for each exam sitting.
```{r eval=FALSE}
download.files.data %>% download.files()
```
